---
title: "Forest Inventory Report"
subtitle: "Automatic report"
author: "r-lidar inc."
date: "`r {Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%B %d, %Y')}`"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    fig_caption: true
    highlight: tango
    includes:
      in_header: preamble.tex
geometry: "left=4cm,right=2cm,top=2cm,bottom=2cm" # Added extra left margin for the band
params:
  idir: "/home/jr/Documents/r-lidar/clients/fsinvestor/Zambia/JasonFarm/qsm/csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center',
  fig.pos = 'H',
  dpi = 400
)

# Load necessary libraries
library(ggplot2)
library(arbor)
library(data.table)
library(sf)
library(kableExtra)
library(patchwork)
library(scales)
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r}
# Define a modern, publication-ready theme
brand_color <- "#1C6B3C" # Matching the LaTeX color

report_theme = theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 12, face = "bold", margin = margin(b = 8)),
    plot.subtitle = element_text(size = 10, color = "grey60", face = "italic"),
    axis.title = element_text(size = 11, face = "bold"),
    axis.text = element_text(size = 10, color = "grey40"),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.5),
    legend.position = "none", # Removed legend as it wasn't used in plots
    plot.background = element_rect(fill = "white", color = NA),
    aspect.ratio = 1
  )

# --- DATA LOADING ---
idir = params$idir

# Check if directory exists to prevent crash if running elsewhere
if(!dir.exists(idir)) stop("Invalid directory")

files <- list.files(idir, full.names = TRUE)

ans <- lapply(files, function(f) 
{
  qsm  <- qsm_read(f)
  name <- tools::file_path_sans_ext(basename(f))
  res  <- qsm_stats(qsm)
  
  root <- qsm[parent_ID == 0]
  # Handle potential empty QSMs safely
  if(nrow(root) > 0){
    root <- st_as_sf(st_sfc(st_point(c(root$startX[1], root$startY[1]))))
    root$name      <- name
    root$volume    <- round(res$stats_global$V, 3)
    root$height    <- round(res$stats_global$H, 2)
    root$dbh       <- res$stats_global$DBH
    root$filename  <- f
    return(root)
  } else {
    return(NULL)
  }
})

ans <- do.call(rbind, ans)

Vtot  <- sum(ans$volume, na.rm = TRUE)
Ntree <- nrow(ans)
idir  <- params$idir
files <- list.files(idir, full.names = TRUE)

ans <- lapply(files, function(f) {

  qsm  <- qsm_read(f)
  name <- tools::file_path_sans_ext(basename(f))
  res  <- qsm_stats(qsm)

  root <- qsm[parent_ID == 0]
  root <- st_as_sf(st_sfc(st_point(c(root$startX, root$startY))))

  root$name     <- name
  root$volume   <- round(res$stats_global$V, 3)
  root$height   <- round(res$stats_global$H, 2)
  root$dbh      <- res$stats_global$DBH
  root$filename <- f

  root
})

ans <- do.call(rbind, ans)
st_crs(ans) <- "EPSG:32735"

Vtot  <- sum(ans$volume)
Ntree <- nrow(ans)
area  <- st_area(st_as_sfc(st_bbox(ans)))
area  <- round(area / 50) * 50
ba    <- sum((ans$dbh/2)^2*pi)
```

# Location

```{r}
# World map
world <- ne_countries(scale = "medium", returnclass = "sf")
rivers <- ne_download(
  scale = 10, type = "rivers_lake_centerlines",
  category = "physical", returnclass = "sf"
)
cities <- ne_download(
  scale = 10, type = "populated_places",
  category = "cultural", returnclass = "sf"
)

bb = st_as_sfc(st_bbox(ans))

centroids <- ans |>
st_transform(3857) |>
st_centroid() |>
st_transform(4326)

bb <- st_bbox(centroids)
pad <- 2
bb_local <- st_bbox(c(
bb["xmin"] - pad,
bb["ymin"] - pad,
bb["xmax"] + pad,
bb["ymax"] + pad
), crs = st_crs(4326))

pad <- 45
bb_global <- st_bbox(c(
bb["xmin"] - pad,
bb["ymin"] - pad,
bb["xmax"] + pad,
bb["ymax"] + pad
), crs = st_crs(4326))

# ---- Local zoom ----
p_local <- ggplot() +
  geom_sf(data = world, fill = "grey98", color = "grey80", linewidth = 0.2) +
  geom_sf(data = rivers, color = "steelblue", linewidth = 0.4, alpha = 0.7) +
  geom_sf(data = cities, color = "black", size = 1) +
  geom_sf_text(
    data = cities,
    aes(label = NAME_EN),
    size = 2,
    check_overlap = TRUE
  ) +
  geom_sf(data = centroids, color = "red", size = 3) +
  coord_sf(
    xlim = c(bb_local["xmin"], bb_local["xmax"]),
    ylim = c(bb_local["ymin"], bb_local["ymax"]),
    expand = FALSE
  ) +
  theme_minimal()


# Plot world map + centroids
p_global <- ggplot() +
geom_sf(data = world, fill = "grey95", color = "grey70", linewidth = 0.2) +
geom_sf(data = centroids, color = "red", size = 2) +
geom_sf(data = st_as_sfc(bb_local), color = "black", size = 1, alpha = 0) +
  coord_sf(
    xlim = c(bb_global["xmin"], bb_global["xmax"]),
    ylim = c(bb_global["ymin"], bb_global["ymax"]),
    expand = FALSE
  ) +
  xlab("") + ylab("") +
theme_minimal()

p_global | p_local
```

# Global Statistics

This report summarizes the QSM processing results. The dataset consists of `r Ntree` trees with a cumulative biological volume of `r round(Vtot, 2)` m^3^.


```{r}
summary_df <- data.frame(
  Metric = c("Trees",
             "Area ",  
             "Density",
             "Basal Area",
             "Basal Area",
             "Total Volume",
             "Average Height", 
             "Average dbh"),
  Value = c(
    as.character(Ntree), 
    paste(round(area, 1), "m$^2$"),
    paste(round(Ntree/area*10000, 2), "tree/ha"),
    paste(round(ba, 2), "m$^2$"),
    paste(round(ba/area*10000, 2), "m$^2/ha$"),
    paste(round(Vtot, 2), "m$^3$"),
    paste(round(mean(ans$height, na.rm=T), 2), "m"),
    paste(round(mean(ans$dbh, na.rm=T)*100, 1), "cm")
  ),
  Description = c(
    "Count of individual tree", 
    "Area of the plot",
    "Number of tree per hectare", 
    "Sum of the cross-sectional area at breast height",
    "Basal area per hectare",
    "Sum of cylindrical volume (Stem + Branches)",
    "Mean vertical height of the stand",
    "Mean diameter at breast height (1.3 m)"
  )
)

kbl(summary_df, booktabs = T, escape = F, align = c('l', 'l', 'l')) %>%
  kable_styling(latex_options = c("striped", "hold_position"), position = "center", font_size = 10) %>%
  row_spec(0, bold = T, color = "white", background = brand_color) %>%
  column_spec(1, bold = T, width = "4cm") %>%
  column_spec(3, italic = T)
```

\pagebreak

# Allometrics

The following figures illustrate the relationships between tree metrics derived from the Quantitative Structure Models.

```{r}
# Plot 1: Volume vs Diameter
p1 <- ggplot(ans, aes(x = dbh * 100, y = volume)) +
geom_point(col = "darkgreen", size = 2, alpha = 0.7) +
labs(
title = "Tree volume vs diameter",
subtitle = "Volume estimated using QSM",
x = "Diameter (cm)",
y = expression(Volume~(m^3))
) +
report_theme


# Plot 2: Height vs Diameter
p2 <- ggplot(ans, aes(x = dbh * 100, y = height)) +
geom_point(col = "darkgreen", size = 2, alpha = 0.7) +
labs(
title = "Tree height vs diameter",
subtitle = "Height estimated using QSM",
x = "Diameter (cm)",
y = "Height (m)"
) +
report_theme


# Display side by side
p1 | p2
```

# Distributions


```{r}
p1 = ggplot(ans, aes(x = dbh * 100)) +
  geom_histogram(bins = 30, fill = "darkgreen", col = "black", alpha = 0.7) +
  labs(
    title = "Distribution of tree diameter",
    subtitle = "Estimated by circle fitting",
    x = "Diameter (cm)"
  ) +
  report_theme

p2 = ggplot(ans, aes(x = height)) +
  geom_histogram(bins = 30, fill = "darkgreen", col = "black", alpha = 0.7) +
  labs(
    title = "Distribution of tree height",
    subtitle = "Estimated using QSM",
    x = "Height (m)"
  ) +
  report_theme

# Display side by side
p1 | p2
```

\pagebreak

# Contributions

```{r}
# --- Data Preparation ---
V <- sort(ans$volume, decreasing = TRUE)           # Sort volumes ascending
Vcum <- cumsum(V)               # Cumulative volume
Vtot <- sum(V)
Vp <- Vcum / Vtot * 100         # Cumulative percentage
Ntree <- length(V)
data <- data.frame(Tree = 1:Ntree, V, Vcum, Vp)

# --- Key Values ---
# Index where cumulative volume reaches 50%
half_vol_idx <- which(Vp >= 50)[1]
# Actual cumulative percentage at that tree
half_vol_value <- Vp[half_vol_idx]
```

The `r half_vol_idx` (`r round((half_vol_idx)/Ntree*100)`%)  biggest trees represent 50% of the volume. The `r Ntree - half_vol_idx` (`r round((Ntree - half_vol_idx)/Ntree*100)`%) smallest trees represent 50% of the volume.


```{r fig.height=4}
p1 <- ggplot(data, aes(x = Tree, y = Vp)) +
  geom_line(color = "darkgreen", size = 1.2) +
  geom_point(color = "forestgreen", size = 2) +
  geom_hline(yintercept = 50, linetype = "dashed", size = 0.5) +
  geom_vline(xintercept = half_vol_idx, linetype = "dashed", size = 0.5) +
  geom_text(
    aes(x = half_vol_idx + 1, y = 98,
        label = paste0(Ntree - half_vol_idx)),
    hjust = 0,
    size = 3.5, # smaller
    fontface = "plain", # not bold
    family = "sans" # vector-friendly font
  ) +
  geom_text(
    aes(x = 0, y = 98,
        label = paste0(half_vol_idx)),
    hjust = 0,
    size = 3.5,
    fontface = "plain",
    family = "sans"
  ) +
  labs(
    title = "Cumulative Volume of Trees",
    subtitle = "Cumulative percentage of total volume with trees order by volume",
    x = "Tree",
    y = "Cumulative Volume (%)"
  ) +
  report_theme

p1
```

# Tree map


```{r}
tree_map = st_buffer(ans, dist = ans$dbh)
ggplot(tree_map) +
  geom_sf(aes(fill = volume), alpha = 0.6, size = 0.2) +
  scale_fill_viridis_c(option = "plasma", name = "Volume (m\U00B3)") +
  labs(
    title = "Tree Map",
    subtitle = "Each tree is represented to 2x scale",
    caption = ""
  ) +
  report_theme + 
theme(
  aspect.ratio = NULL,
  legend.position = c(1.2, 1), # x and y in [0,1], top-right
  legend.justification = c("right", "top"),
  legend.background = element_rect(fill = "white", color = "black", size = 0.25, linetype = "solid"),
  legend.key = element_rect(fill = "white"), # key background
  legend.key.size = unit(0.25, "cm"), # smaller color boxes
  legend.title = element_text(size = 8), # smaller title
  legend.text = element_text(size = 8) # smaller labels
  ) +
  coord_sf()
```

