---
title: "Forest Inventory Report"
subtitle: "Automatic report"
author: "By r-lidar inc."
date: "`r {Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%B %d, %Y')}`"
output:
  html_document:
    theme: default           # minimal base theme; custom.css overrides its look
    css: custom.css
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: false
    code_folding: hide
    fig_caption: true
    self_contained: true
params:
  input: "/home/jr/R packages/r-lidar-lab/arbor/inst/extdata/oak-plantation.qsf"
  output: "/home/jr/R packages/r-lidar-lab/arbor/inst/extdata/oak-plantation.html"
  target_epsg: 32619
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center'
)

library(ggplot2)
library(arbor)
library(data.table)
library(sf)
library(scales)
library(leaflet)
library(plotly)
library(DT)
library(crosstalk)
library(htmltools)
library(jsonlite)

# --- ACCURATE DATUM TRANSFORMS, WITH A SAFETY NET -----------------------
# NAD27 (the datum under EPSG:32098) needs an NTv2 grid file for an accurate
# shift to WGS84; that grid isn't bundled with sf, only fetchable from the
# PROJ CDN. If this machine can't actually reach that CDN, telling PROJ to
# prefer the grid-based pipeline (sf_proj_network(TRUE)) backfires: PROJ
# still picks that pipeline, the grid never downloads, and every point
# comes back degenerate instead of falling back gracefully - which is what
# pushes everything to the pole in a Lambert Conformal Conic projection.
# So: try the accurate transform, but verify the result actually lands in
# Quebec before trusting it; if not, disable network mode and redo it with
# whatever grids/approximation are available locally instead.
qc_bbox_wgs84 <- c(xmin = -80, xmax = -55, ymin = 44, ymax = 63)

safe_transform_to_wgs84 <- function(x) {
  transform_and_check <- function(x) {
    xy <- sf::st_coordinates(sf::st_transform(x, 4326))
    ok <- xy[, 1] >= qc_bbox_wgs84["xmin"] & xy[, 1] <= qc_bbox_wgs84["xmax"] &
          xy[, 2] >= qc_bbox_wgs84["ymin"] & xy[, 2] <= qc_bbox_wgs84["ymax"]
    list(xy = xy, ok = all(ok))
  }

  tryCatch(sf::sf_proj_network(TRUE), error = function(e) invisible(NULL))
  res <- transform_and_check(x)

  if (!res$ok) {
    warning(
      "Network datum grid produced points outside Quebec - probably ",
      "unreachable from this machine. Falling back to the local ",
      "(less accurate, ~10-100 m) NAD27->WGS84 approximation.",
      call. = FALSE
    )
    sf::sf_proj_network(FALSE)
    res <- transform_and_check(x)
    if (!res$ok)
      warning(
        "Points are still outside Quebec even with the local fallback. ",
        "This points to the source X/Y values themselves not being real ",
        "EPSG:32098 (NAD27 / Quebec Lambert) coordinates - check the raw ",
        "X_root/Y_root against the CRS's valid range (easting roughly ",
        "-900,000 to 900,000 m, northing roughly 110,000 to 2,112,000 m).",
        call. = FALSE
      )
  }

  res$xy
}
```

```{r brand-palette}
# Keep this in sync with the :root variables in custom.css
brand_green   <- "#1C6B3C"
brand_green_d <- "#123F23"
accent_amber  <- "#C77B3B"   # used for click/selection highlight
ink           <- "#2B2B28"
mid_gray      <- "#6B6963"

report_theme <- theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(size = 13, face = "bold", margin = margin(b = 6)),
    plot.subtitle = element_text(size = 10, color = mid_gray, face = "italic"),
    axis.title    = element_text(size = 11, face = "bold"),
    axis.text     = element_text(size = 10, color = mid_gray),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    legend.position  = "none",
    plot.background  = element_rect(fill = "white", color = NA)
  )
```

```{r read-qsm}
input <- params$input
output_file <- params$output
target_epsg <- params$target_epsg

if (file.info(input)$isdir)
{
  f <- list.files(input, pattern = "qsf", recursive = T, full.names = T)
  if(length(f) == 0) stop("No .qsf file detected in this directory")
  ans <- lapply(f, arbor::qsf_read)
  qsf <- do.call(c, ans)
  qsf <- arbor:::as_qsf(qsf)
} else if (file.exists(input)) {
  qsf <- qsf_read(input)
} else {
   stop("Invalid file or directory")
}

ans <- qsf_treemap(qsf)

has_crs <- !is.na(st_crs(ans))

if (target_epsg > 0) 
{
  if (!has_crs)
    st_crs(ans) <- target_epsg
  else
    ans <- st_transform(ans, target_epsg)
}

has_crs <- !is.na(st_crs(ans))

ogpkg = paste0(tools::file_path_sans_ext(output_file), ".gpkg")
st_write(ans, ogpkg, quiet = TRUE, append = FALSE)

ans <- dplyr::rename(ans, name = Name, volume = V, height = H, dbh = DBH)
  
Vtot  <- sum(ans$volume)
Ntree <- nrow(ans)
area  <- st_area(st_as_sfc(st_bbox(ans)))
area  <- round(area / 50) * 50
ba    <- sum((ans$dbh / 2)^2 * pi)
```

```{r shared-data}
# --- ONE shared dataset feeds every interactive widget below ---------------
# Clicking a point in any plot, a marker on the tree map, or a row in the
# table highlights the *same tree everywhere else, because they all point
# back to this single crosstalk group ("trees"). The volume slider and the
# status checkboxes at the top of the report (see "Filters" below) act on
# this same group too, so they subset every linked plot, map and table in
# sync -- including the Global Statistics cards, via a small JS listener.

ans_df <- st_drop_geometry(ans)
ans_df$dbh_cm <- round(ans_df$dbh * 100, 1)

if (has_crs) {
  coords <- safe_transform_to_wgs84(ans)
  ans_df$lon <- coords[, 1]
  ans_df$lat <- coords[, 2]
}

# Normalize the QSM "Message" flag once: missing values become "OK", and we
# fix a stable factor level order + color palette (OK is always brand green)
# that every chart/table/checkbox below reuses. The column stays named
# "Message" in code -- only the labels shown to the reader say "Status".
ans_df$Message <- ifelse(is.na(ans_df$Message), "OK", as.character(ans_df$Message))
status_levels  <- c("OK", sort(setdiff(unique(ans_df$Message), "OK")))
status_colors  <- setNames(
  c(brand_green, scales::hue_pal()(max(length(status_levels) - 1, 0))),
  status_levels
)
ans_df$Message <- factor(ans_df$Message, levels = status_levels)

sd_trees <- SharedData$new(ans_df, key = ~name, group = "trees")
```

# Location

Interactive map of the plot: scroll to zoom, drag to pan.

```{r world-map, fig.height=5}
if (has_crs) {

  hull <- st_convex_hull(st_union(st_zm(ans)))
  hull <- st_as_sf(hull)
  hull <- st_transform(hull, 4326)

  # 1. Compute the centroid point for the marker
  centroid <- st_centroid(hull)
  coords <- st_coordinates(centroid)
  lng <- coords[1]
  lat <- coords[2]

  leaflet(width = "100%", height = 480) %>%
    # 2. Set custom world view zoom (zoom level 2 gives a macro view)
    setView(lng = lng, lat = lat, zoom = 2) %>% 
    addProviderTiles(providers$Esri.WorldTopoMap, group = "Topo (Esri)") %>%
    addProviderTiles(providers$Esri.WorldImagery, group = "Satellite (Esri)") %>%
    addPolygons(
      data = hull,
      weight = 2,
      color = "#ffffff",
      fillOpacity = 0.6,
      highlightOptions = highlightOptions(
        weight = 3,
        color = "#00ffff",
        fillOpacity = 0.8,
        bringToFront = TRUE
      )
    ) %>%
    # 3. Add a location marker at the polygon center
    addMarkers(
      data = centroid,
      popup = "Location"
    ) %>%
    addLayersControl(
      baseGroups = c("Topo (Esri)", "Satellite (Esri)"),
      options = layersControlOptions(collapsed = TRUE)
    ) %>%
    addScaleBar(position = "bottomleft")

} else {

  div(class = "callout",
      strong("Cannot display data location \u2014 "),
      "the point cloud/QSM data has no coordinate reference system (CRS) attached."
  )
}
```

# Filters

Use the controls below to focus the report on a subset of trees. Both
controls act on the **same set of trees** and immediately update every
chart, map and table below.

```{r filters-slider}
vol_range <- range(ans_df$volume, na.rm = TRUE)
vol_step  <- signif(diff(vol_range) / 500, 2)
if (!is.finite(vol_step) || vol_step <= 0) vol_step <- 0.001

filter_slider(
  id = "volume_filter",
  label = "Filter by tree volume (m³)",
  sharedData = sd_trees,
  column = ~volume,
  step = vol_step,
  width = "100%"
)
```

```{r filters-status}
filter_checkbox(
  id = "status_filter",
  label = "Filter by status",
  sharedData = sd_trees,
  group = ~Message,
  inline = TRUE
)
```

# Global Statistics

This report summarizes the QSM processing results. The dataset consists of
`r Ntree` trees with a cumulative biological volume of `r round(Vtot, 2)` m^3^.

```{r stat-cards}
stat <- function(label, value, desc, id = NULL) {
  div(class = "stat-card",
      div(class = "stat-label", label),
      div(class = "stat-value", if (is.null(id)) value else tags$span(id = id, value)),
      div(class = "stat-desc", desc)
  )
}

tagList(
  div(class = "stat-grid",
    stat("Trees", Ntree, "Count of individual trees", id = "stat-trees"),
    stat("Area", paste(round(area, 1), "m\u00B2"),
         "Area of the plot (fixed \u2014 the surveyed footprint doesn't shrink when you filter trees)"),
    stat("Density", paste(round(Ntree / area * 10000, 2), "/ ha"), "Trees per hectare", id = "stat-density"),
    stat("Basal area", paste(round(ba, 2), "m\u00B2"), "Sum of cross-sectional area at breast height", id = "stat-basal-area"),
    stat("Basal area / ha", paste(round(ba / area * 10000, 2), "m\u00B2/ha"), "Basal area per hectare", id = "stat-basal-area-ha"),
    stat("Total volume", paste(round(Vtot, 2), "m\u00B3"), "Sum of cylindrical volume (stem + branches)", id = "stat-total-volume"),
    stat("Avg. height", paste(round(mean(ans_df$height, na.rm = TRUE), 2), "m"), "Mean vertical height of the stand", id = "stat-avg-height"),
    stat("Avg. DBH", paste(round(mean(ans_df$dbh_cm, na.rm = TRUE), 1), "cm"), "Mean diameter at breast height (1.3 m)", id = "stat-avg-dbh")
  ),
  tags$script(
    id = "trees-data-json", type = "application/json",
    HTML(jsonlite::toJSON(ans_df[, c("name", "volume", "height", "dbh")], digits = 8))
  ),
  tags$script(HTML(sprintf("
    (function() {
      const totalArea = %f;
      const dataEl = document.getElementById('trees-data-json');
      const allTrees = JSON.parse(dataEl.textContent);

      function computeStats(subset) {
        let n = subset.length, vol = 0, ba = 0, height = 0, dbhCm = 0;
        subset.forEach(function(t) {
          vol += t.volume;
          ba  += Math.PI * Math.pow(t.dbh / 2, 2);
          height += t.height;
          dbhCm += t.dbh * 100;
        });
        return {
          n: n, volume: vol, ba: ba,
          avgHeight: n > 0 ? height / n : 0,
          avgDbh: n > 0 ? dbhCm / n : 0
        };
      }

      function render(subset) {
        const s = computeStats(subset);
        document.getElementById('stat-trees').textContent = s.n;
        document.getElementById('stat-density').textContent = (s.n / totalArea * 10000).toFixed(2) + ' / ha';
        document.getElementById('stat-basal-area').textContent = s.ba.toFixed(2) + ' m\u00B2';
        document.getElementById('stat-basal-area-ha').textContent = (s.ba / totalArea * 10000).toFixed(2) + ' m\u00B2/ha';
        document.getElementById('stat-total-volume').textContent = s.volume.toFixed(2) + ' m\u00B3';
        document.getElementById('stat-avg-height').textContent = s.avgHeight.toFixed(2) + ' m';
        document.getElementById('stat-avg-dbh').textContent = s.avgDbh.toFixed(1) + ' cm';
      }

      function applyFilterKeys(keys) {
        if (keys === null || typeof keys === 'undefined') {
          render(allTrees);
        } else {
          const keySet = {};
          keys.forEach(function(k) { keySet[k] = true; });
          render(allTrees.filter(function(t) { return keySet[t.name]; }));
        }
      }

      // Crosstalk combines the volume slider and the status checkboxes into
      // one intersected 'filter' variable per group, so a single listener
      // here stays correct no matter how many filter widgets are active.
      function bindWhenReady() {
        if (window.crosstalk && crosstalk.group) {
          const filterVar = crosstalk.group('trees').var('filter');
          filterVar.on('change', function(e) { applyFilterKeys(e.value); });
          applyFilterKeys(filterVar.get());
        } else {
          setTimeout(bindWhenReady, 50);
        }
      }
      bindWhenReady();
    })();
  ", as.numeric(area))))
)
```

# Carbon Estimate

Total stand volume can be converted into an approximate carbon stock using a
conversion factor (Mg C per m&sup3; of wood volume), which depends on
species, wood density, and the carbon fraction of biomass. That factor isn't
fixed for this dataset yet, use the slider below to explore the estimate
under different assumptions, and swap in the correct factor for your
species/region once you have it.

```{r carbon-estimate}
tagList(
  div(class = "callout",
    tags$label(
      `for` = "carbon-slider",
      style = "font-weight:600; display:block; margin-bottom:8px;",
      "Conversion factor: ",
      tags$span(id = "carbon-factor-display", "0.50"),
      " Mg C / m\u00B3"
    ),
    tags$input(
      id = "carbon-slider", type = "range",
      min = "0.05", max = "1", step = "0.01", value = "0.5",
      style = "width:100%; accent-color:#1C6B3C;"
    )
  ),
  div(class = "stat-grid",
    stat("Total carbon", tags$span(id = "total-carbon-value", "-"), "Estimated stand carbon stock"),
    stat("Carbon / ha", tags$span(id = "carbon-per-ha-value", "-"), "Estimated carbon density"),
    stat("Total volume", paste(round(Vtot, 2), "m\u00B3"), "Reference: total QSM volume (fixed)")
  ),
  tags$script(HTML(sprintf("
    (function() {
      const vtot = %f;
      const area = %f;
      const slider = document.getElementById('carbon-slider');
      const factorDisplay = document.getElementById('carbon-factor-display');
      const totalEl = document.getElementById('total-carbon-value');
      const perHaEl = document.getElementById('carbon-per-ha-value');

      function update() {
        const factor = parseFloat(slider.value);
        factorDisplay.textContent = factor.toFixed(2);
        const total = vtot * factor;
        const perHa = (total / area) * 10000;
        totalEl.textContent = total.toFixed(2) + ' Mg';
        perHaEl.textContent = perHa.toFixed(2) + ' Mg/ha';
      }

      slider.addEventListener('input', update);
      update();
    })();
  ", Vtot, as.numeric(area))))
)
```

# QSM Quality

Each tree carries a status flag from the QSM reconstruction process,
indicating whether the model completed cleanly or with a warning. The chart
below shows what share of the **stand's total volume** falls under each
status — this matters more than a simple tree count, since a warning on one
large tree can carry more volume risk than several small ones. Trees with no
status recorded are grouped under "OK".

```{r qsm-quality}
qual <- ans_df

qual_summary <- aggregate(volume ~ Message, data = qual, FUN = sum)
qual_summary$n <- as.integer(table(qual$Message)[as.character(qual_summary$Message)])
qual_summary$pct <- round(qual_summary$volume / sum(qual_summary$volume) * 100, 1)
qual_summary <- qual_summary[order(-qual_summary$volume), ]

qual_colors <- status_colors[as.character(qual_summary$Message)]

plot_ly(
  qual_summary,
  labels = ~Message,
  values = ~volume,
  type = "pie",
  hole = 0.55,
  rotation = 90, # <-- Add this to rotate the starting point to the right
  textinfo = "label+percent",
  customdata = ~n,
  hovertemplate = paste0(
    "<b>%{label}</b><br>",
    "%{value:.2f} m\u00B3 (%{percent})<br>",
    "%{customdata} trees<extra></extra>"
  ),
  marker = list(colors = qual_colors, line = list(color = "white", width = 1.5))
) %>%
  layout(
    title = list(text = "Share of total volume by status", font = list(size = 13)),
    font = list(family = "IBM Plex Sans, sans-serif"),
    showlegend = TRUE,
    annotations = list(list(
      text = paste0(round(Vtot, 1), " m\u00B3"),
      x = 0.5, y = 0.5, font = list(size = 14, color = ink), showarrow = FALSE
    ))
  )
```

```{r qsm-quality-table}
datatable(
  qual_summary[, c("Message", "n", "volume", "pct")],
  rownames = FALSE,
  colnames = c("Status", "Trees", "Volume (m\u00B3)", "% of total volume"),
  options = list(dom = "t", pageLength = nrow(qual_summary), ordering = FALSE),
  class = "display compact stripe"
) %>%
  formatRound(columns = "volume", digits = 2)
```

# Allometrics {.tabset}

The tabs below show the relationships between tree metrics derived from the
Quantitative Structure Models. **Click a point** in any tab (or a row in the
*Data table* tab) to highlight that tree everywhere on this page, including
the tree map further down. Double-click a plot to clear the selection. The
volume and status filters near the top of this report also apply here.

## Volume vs Diameter

```{r allometry-vol}
p1 <- ggplot(sd_trees, aes(x = dbh_cm, y = volume, color = Message,
                           text = paste0("Tree: ", name, "<br>",
                                         "Status: ", as.character(Message), "<br>",
                                         dbh_cm, " cm \u2014 ", volume, " m\u00B3"))) +
  geom_point(size = 2, alpha = 0.75) +
  scale_color_manual(values = status_colors) +
  labs(title = "Tree volume vs diameter", subtitle = "Volume estimated using QSM",
       x = "Diameter (cm)", y = "Volume (m\u00B3)", color = "Status") +
  report_theme

ggplotly(p1, tooltip = "text", height = 420) %>%
  highlight(on = "plotly_click", off = "plotly_doubleclick",
            color = accent_amber, opacityDim = 0.2) %>%
  layout(
    font = list(family = "IBM Plex Sans, sans-serif"),
    showlegend = FALSE
  )
```

## Height vs Diameter

```{r allometry-height}
p2 <- ggplot(sd_trees, aes(x = dbh_cm, y = height, color = Message,
                           text = paste0("Tree: ", name, "<br>",
                                         "Status: ", as.character(Message), "<br>",
                                         dbh_cm, " cm \u2014 ", height, " m"))) +
  geom_point(size = 2, alpha = 0.75) +
  scale_color_manual(values = status_colors) +
  labs(title = "Tree height vs diameter", subtitle = "Height estimated using QSM",
       x = "Diameter (cm)", y = "Height (m)", color = "Status") +
  report_theme

ggplotly(p2, tooltip = "text", height = 420) %>%
  highlight(on = "plotly_click", off = "plotly_doubleclick",
            color = accent_amber, opacityDim = 0.2) %>%
  layout(
    font = list(family = "IBM Plex Sans, sans-serif"),
    showlegend = FALSE
  )
```

## Data table

The table below can be downloaded as CSV or Excel using the buttons above it
(the export respects any column filters you've applied).

```{r allometry-table}
datatable(
  sd_trees,
  rownames = FALSE,
  filter = "top", # Will automatically use a multi-select dropdown for factors
  extensions = "Buttons",
  colnames = c(
    "Tree" = "name", 
    "Status" = "Message", 
    "Volume (m\u00B3)" = "volume",
    "Height (m)" = "height", 
    "DBH (cm)" = "dbh_cm"
  ),
  options = list(
    pageLength = 18, 
    dom = "Bftip",
    buttons = list(
      list(extend = "csv",   text = "Download CSV",   filename = "tree_inventory", exportOptions = list(modifier = list(search = "applied"))),
      list(extend = "excel", text = "Download Excel", filename = "tree_inventory", exportOptions = list(modifier = list(search = "applied")))
    ),
    columnDefs = list(
      list(
        visible = FALSE, 
        targets = which(names(ans_df) %in% c("dbh", "filename", "lon", "lat")) - 1
      )
    )
  ),
  class = "display compact stripe"
)
```

# Distributions {.tabset}

## Diameter

```{r distribution-diam}
p1 <- ggplot(sd_trees, aes(x = dbh_cm,
                          text = after_stat(paste0(count, " trees, ", round(x, 1), " cm bin")))) +
  geom_histogram(bins = 30, fill = brand_green, col = "white", alpha = 0.9) +
  labs(title = "Distribution of tree diameter", subtitle = "Estimated by circle fitting",
       x = "Diameter (cm)", y = "Count") +
  report_theme

ggplotly(p1, tooltip = "text", height = 380) %>%
  layout(font = list(family = "IBM Plex Sans, sans-serif"))
```

## Height

```{r distribution-height}
p2 <- ggplot(sd_trees, aes(x = height,
                          text = after_stat(paste0(count, " trees, ", round(x, 1), " m bin")))) +
  geom_histogram(bins = 30, fill = brand_green, col = "white", alpha = 0.9) +
  labs(title = "Distribution of tree height", subtitle = "Estimated using QSM",
       x = "Height (m)", y = "Count") +
  report_theme

ggplotly(p2, tooltip = "text", height = 380) %>%
  layout(font = list(family = "IBM Plex Sans, sans-serif"))
```

# Contributions

```{r contribution}
data <- ans_df[order(-ans_df$volume), ]
data$Tree <- seq_len(nrow(data))
data$Vcum <- cumsum(data$volume)
data$Vp   <- data$Vcum / sum(data$volume) * 100

half_vol_idx   <- which(data$Vp >= 50)[1]
half_vol_value <- data$Vp[half_vol_idx]

sd_contrib <- SharedData$new(data, key = ~name, group = "trees")
```

<div class="callout">
The **`r half_vol_idx`** (`r round(half_vol_idx / Ntree * 100)`%) biggest trees represent 50% of the volume.
The **`r Ntree - half_vol_idx`** (`r round((Ntree - half_vol_idx) / Ntree * 100)`%) smallest trees represent the other 50%.
</div>

Click a point below to see which tree it is (linked to the table and plots above).

```{r contribution2}
p1 <- ggplot(sd_contrib, aes(x = Tree, y = Vp,
                              text = paste0("Tree: ", name, "<br>rank ", Tree, " \u2014 ", round(Vp, 1), "% cumulative"))) +
  geom_line(color = brand_green, linewidth = 1, group = 1) +
  geom_point(color = brand_green, size = 1.6) +
  geom_hline(yintercept = 50, linetype = "dashed", color = mid_gray) +
  geom_vline(xintercept = half_vol_idx, linetype = "dashed", color = mid_gray) +
  labs(title = "Cumulative volume of trees",
       subtitle = "Trees ordered by volume, largest first",
       x = "Tree rank", y = "Cumulative volume (%)") +
  report_theme

ggplotly(p1, tooltip = "text", height = 420) %>%
  highlight(on = "plotly_click", off = "plotly_doubleclick",
            color = accent_amber, opacityDim = 0.2) %>%
  layout(font = list(family = "IBM Plex Sans, sans-serif"))
```

# Tree map

Each tree's diameter footprint is drawn at 2x scale and colored by volume. Scroll to zoom; click a tree for details — the selection stays in sync with the charts and table above.
```{r tree-map, fig.height=5}
local_xy <- st_coordinates(ans)
ans_df$x_local <- local_xy[, 1]
ans_df$y_local <- local_xy[, 2]

# --- Pre-compute the three "color by" options -----------------------------
# Volume / Height are continuous -> Viridis colorscale + colorbar.
# Status is discrete -> a fixed hex color per tree, colorbar hidden (uses
# the same status_colors palette as every other chart in this report).
# All three live on ONE trace; the buttons below just restyle/relayout it,
# so the crosstalk selection group never has to be rebuilt.
ans_df$msg_color <- unname(status_colors[as.character(ans_df$Message)])

ans_df$hover_text <- paste0(
  "<b>Tree: ", ans_df$name, "</b><br>",
  "Volume: ", ans_df$volume, " m\u00B3<br>",
  "Height: ", ans_df$height, " m<br>",
  "DBH: ", ans_df$dbh_cm, " cm<br>",
  "Status: ", as.character(ans_df$Message)
)

sd_local <- SharedData$new(ans_df, key = ~name, group = "trees")

vol_range    <- range(ans_df$volume, na.rm = TRUE)
height_range <- range(ans_df$height, na.rm = TRUE)
size_ref     <- 2 * max(ans_df$volume, na.rm = TRUE) / (26^2)

p <- plot_ly(
  data = sd_local,
  x = ~x_local, y = ~y_local,
  type = "scatter", mode = "markers",
  text = ~hover_text, hoverinfo = "text",
  height = 480,
  marker = list(
    color = ~volume,
    colorscale = "Viridis",
    cmin = vol_range[1], cmax = vol_range[2],
    showscale = TRUE,
    colorbar = list(title = list(text = "Volume (m\u00B3)")),
    size = ~volume, sizemode = "area",
    sizeref = size_ref, sizemin = 4,
    line = list(color = "white", width = 0.4),
    opacity = 0.75
  )
) %>%
  highlight(
    on = "plotly_click", off = "plotly_doubleclick",
    color = accent_amber, opacityDim = 0.2
  ) %>%
  layout(
    title = list(text = "Tree map", font = list(size = 13)),
    xaxis = list(title = "X"),
    yaxis = list(title = "Y"),
    font = list(family = "IBM Plex Sans, sans-serif"),
    updatemenus = list(
      list(
        type = "buttons", direction = "right", active = 0,
        x = 1, xanchor = "right",
        y = 1.15, yanchor = "top",
        pad = list(r = 4, t = 4),
        buttons = list(
          list(
            label = "Volume",
            method = "update",
            args = list(
              list(
                "marker.color"               = list(ans_df$volume),
                "marker.colorscale"          = list("Viridis"),
                "marker.cmin"                = list(vol_range[1]),
                "marker.cmax"                = list(vol_range[2]),
                "marker.showscale"           = list(TRUE),
                "marker.colorbar.title.text" = list("Volume (m\u00B3)")
              ),
              list(title = list(text = "Tree map"))
            )
          ),
          list(
            label = "Height",
            method = "update",
            args = list(
              list(
                "marker.color"               = list(ans_df$height),
                "marker.colorscale"          = list("Viridis"),
                "marker.cmin"                = list(height_range[1]),
                "marker.cmax"                = list(height_range[2]),
                "marker.showscale"           = list(TRUE),
                "marker.colorbar.title.text" = list("Height (m)")
              ),
              list(title = list(text = "Tree map"))
            )
          ),
          list(
            label = "Status",
            method = "update",
            args = list(
              list(
                "marker.color"     = list(ans_df$msg_color),
                "marker.showscale" = list(FALSE)
              ),
              list(title = list(text = "Tree map"))
            )
          )
        )
      )
    )
  )

# Render with plotly_build to process HTML correctly
plotly_build(p)
```

<div class="report-footer">
Generated automatically by r-lidar inc. &middot; `r format(Sys.time(), "%Y-%m-%d %H:%M")`
</div>
