---
title: "Forest Inventory Report"
subtitle: "Automatic report"
author: "By r-lidar inc."
date: "`r {Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%B %d, %Y')}`"
output:
  html_document:
    theme: default           # minimal base theme; custom.css overrides its look
    css: custom.css
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: false
    code_folding: hide
    fig_caption: true
    self_contained: true
params:
  input: "/home/jr/R packages/r-lidar-lab/arbor/inst/extdata/oak-plantation.qsf"
  epsg: 32619
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center'
)

library(ggplot2)
library(arbor)
library(data.table)
library(sf)
library(scales)
library(leaflet)
library(plotly)
library(DT)
library(crosstalk)
library(htmltools)

# --- ACCURATE DATUM TRANSFORMS, WITH A SAFETY NET -----------------------
# NAD27 (the datum under EPSG:32098) needs an NTv2 grid file for an accurate
# shift to WGS84; that grid isn't bundled with sf, only fetchable from the
# PROJ CDN. If this machine can't actually reach that CDN, telling PROJ to
# prefer the grid-based pipeline (sf_proj_network(TRUE)) backfires: PROJ
# still picks that pipeline, the grid never downloads, and every point
# comes back degenerate instead of falling back gracefully - which is what
# pushes everything to the pole in a Lambert Conformal Conic projection.
# So: try the accurate transform, but verify the result actually lands in
# Quebec before trusting it; if not, disable network mode and redo it with
# whatever grids/approximation are available locally instead.
qc_bbox_wgs84 <- c(xmin = -80, xmax = -55, ymin = 44, ymax = 63)

safe_transform_to_wgs84 <- function(x) {
  transform_and_check <- function(x) {
    xy <- sf::st_coordinates(sf::st_transform(x, 4326))
    ok <- xy[, 1] >= qc_bbox_wgs84["xmin"] & xy[, 1] <= qc_bbox_wgs84["xmax"] &
          xy[, 2] >= qc_bbox_wgs84["ymin"] & xy[, 2] <= qc_bbox_wgs84["ymax"]
    list(xy = xy, ok = all(ok))
  }

  tryCatch(sf::sf_proj_network(TRUE), error = function(e) invisible(NULL))
  res <- transform_and_check(x)

  if (!res$ok) {
    warning(
      "Network datum grid produced points outside Quebec - probably ",
      "unreachable from this machine. Falling back to the local ",
      "(less accurate, ~10-100 m) NAD27->WGS84 approximation.",
      call. = FALSE
    )
    sf::sf_proj_network(FALSE)
    res <- transform_and_check(x)
    if (!res$ok)
      warning(
        "Points are still outside Quebec even with the local fallback. ",
        "This points to the source X/Y values themselves not being real ",
        "EPSG:32098 (NAD27 / Quebec Lambert) coordinates - check the raw ",
        "X_root/Y_root against the CRS's valid range (easting roughly ",
        "-900,000 to 900,000 m, northing roughly 110,000 to 2,112,000 m).",
        call. = FALSE
      )
  }

  res$xy
}
```

```{r brand-palette}
# Keep this in sync with the :root variables in custom.css
brand_green   <- "#1C6B3C"
brand_green_d <- "#123F23"
accent_amber  <- "#C77B3B"   # used for click/selection highlight
ink           <- "#2B2B28"
mid_gray      <- "#6B6963"

report_theme <- theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(size = 13, face = "bold", margin = margin(b = 6)),
    plot.subtitle = element_text(size = 10, color = mid_gray, face = "italic"),
    axis.title    = element_text(size = 11, face = "bold"),
    axis.text     = element_text(size = 10, color = mid_gray),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    legend.position  = "none",
    plot.background  = element_rect(fill = "white", color = NA)
  )
```

```{r read-qsm}
input <- params$input
epsg <- params$epsg
if (!file.exists(input)) stop("Invalid directory")

qsf <- qsf_read(input)
ans <- qsf_treemap(qsf)

has_crs <- !is.na(st_crs(ans))

if (epsg > 0) 
{
  if (!has_crs)
    st_crs(ans) <- epsg
  else
    ans <- st_transform(ans, epsg)
}

has_crs <- !is.na(st_crs(ans))

st_write(ans, paste0(tools::file_path_sans_ext(input), ".gpkg"), quiet = TRUE)

ans <- dplyr::rename(ans, name = Name, volume = V, height = H, dbh = DBH)
  
Vtot  <- sum(ans$volume)
Ntree <- nrow(ans)
area  <- st_area(st_as_sfc(st_bbox(ans)))
area  <- round(area / 50) * 50
ba    <- sum((ans$dbh / 2)^2 * pi)


```

```{r shared-data}
# --- ONE shared dataset feeds every interactive widget below ---------------
# Clicking a point in any plot, a marker on the tree map, or a row in the
# table highlights the *same tree* everywhere else, because they all point
# back to this single crosstalk group ("trees").

ans_df <- st_drop_geometry(ans)
ans_df$dbh_cm <- round(ans_df$dbh * 100, 1)

if (has_crs) {
  coords <- safe_transform_to_wgs84(ans)
  ans_df$lon <- coords[, 1]
  ans_df$lat <- coords[, 2]
}

sd_trees <- SharedData$new(ans_df, key = ~name, group = "trees")
```

# Location

Interactive map of the plot — scroll to zoom, drag to pan. The inset in the
bottom-right corner shows the wider region for context (click it to expand).

```{r world-map, fig.height=5}
if (has_crs) {

  hull <- st_convex_hull(st_union(st_zm(ans)))
  hull <- st_as_sf(hull)
  hull <- st_transform(hull, 4326)

  # 1. Compute the centroid point for the marker
  centroid <- st_centroid(hull)
  coords <- st_coordinates(centroid)
  lng <- coords[1]
  lat <- coords[2]

  leaflet(width = "100%", height = 480) %>%
    # 2. Set custom world view zoom (zoom level 2 gives a macro view)
    setView(lng = lng, lat = lat, zoom = 2) %>% 
    addProviderTiles(providers$Esri.WorldTopoMap, group = "Topo (Esri)") %>%
    addProviderTiles(providers$Esri.WorldImagery, group = "Satellite (Esri)") %>%
    addPolygons(
      data = hull,
      weight = 2,
      color = "#ffffff",
      fillOpacity = 0.6,
      highlightOptions = highlightOptions(
        weight = 3,
        color = "#00ffff",
        fillOpacity = 0.8,
        bringToFront = TRUE
      )
    ) %>%
    # 3. Add a location marker at the polygon center
    addMarkers(
      data = centroid,
      popup = "Location"
    ) %>%
    addLayersControl(
      baseGroups = c("Topo (Esri)", "Satellite (Esri)"),
      options = layersControlOptions(collapsed = TRUE)
    ) %>%
    addMiniMap(
      tiles = providers$CartoDB.Positron,
      toggleDisplay = TRUE, minimized = TRUE, position = "bottomright"
    ) %>%
    addScaleBar(position = "bottomleft")

} else {

  div(class = "callout",
      strong("Cannot display data location \u2014 "),
      "the point cloud/QSM data has no coordinate reference system (CRS) attached."
  )
}
```

# Global Statistics

This report summarizes the QSM processing results. The dataset consists of
`r Ntree` trees with a cumulative biological volume of `r round(Vtot, 2)` m^3^.

```{r stat-cards}
stat <- function(label, value, desc) {
  div(class = "stat-card",
      div(class = "stat-label", label),
      div(class = "stat-value", value),
      div(class = "stat-desc", desc)
  )
}

div(class = "stat-grid",
  stat("Trees", Ntree, "Count of individual trees"),
  stat("Area", paste(round(area, 1), "m\u00B2"), "Area of the plot"),
  stat("Density", paste(round(Ntree / area * 10000, 2), "/ ha"), "Trees per hectare"),
  stat("Basal area", paste(round(ba, 2), "m\u00B2"), "Sum of cross-sectional area at breast height"),
  stat("Basal area / ha", paste(round(ba / area * 10000, 2), "m\u00B2/ha"), "Basal area per hectare"),
  stat("Total volume", paste(round(Vtot, 2), "m\u00B3"), "Sum of cylindrical volume (stem + branches)"),
  stat("Avg. height", paste(round(mean(ans_df$height, na.rm = TRUE), 2), "m"), "Mean vertical height of the stand"),
  stat("Avg. DBH", paste(round(mean(ans_df$dbh_cm, na.rm = TRUE), 1), "cm"), "Mean diameter at breast height (1.3 m)")
)
```

# Allometrics {.tabset}

The tabs below show the relationships between tree metrics derived from the
Quantitative Structure Models. **Click a point** in any tab (or a row in the
*Data table* tab) to highlight that tree everywhere on this page — including
the tree map further down. Double-click a plot to clear the selection. Use
the slider to filter the trees under study.

```{r allometry-filter}
filter_slider(
  id = "volume_filter",
  label = "Filter by tree volume (m³)",
  sharedData = sd_trees,
  column = ~volume,
  width = "100%"
)
```

## Volume vs Diameter

```{r allometry-vol}
msg_levels <- na.omit(unique(ans_df$Message))
n_msgs <- length(msg_levels)

# Generate colors dynamically for non-NA messages
msg_colors <- setNames(scales::hue_pal()(n_msgs), msg_levels)

p1 <- ggplot(sd_trees, aes(x = dbh_cm, y = volume, color = Message,
                           text = paste0("Tree: ", name, "<br>",
                                         "Message: ", ifelse(is.na(Message), "None", as.character(Message)), "<br>",
                                         dbh_cm, " cm \u2014 ", volume, " m\u00B3"))) +
  geom_point(size = 2, alpha = 0.75) +
  scale_color_manual(
    na.value = brand_green,
    values = msg_colors
  ) +
  labs(title = "Tree volume vs diameter", subtitle = "Volume estimated using QSM",
       x = "Diameter (cm)", y = "Volume (m\u00B3)", color = "Message") +
  report_theme

ggplotly(p1, tooltip = "text", height = 420) %>%
  highlight(on = "plotly_click", off = "plotly_doubleclick",
            color = accent_amber, opacityDim = 0.2) %>%
  layout(
    font = list(family = "IBM Plex Sans, sans-serif"),
    showlegend = FALSE
  )
```

## Height vs Diameter

```{r allometry-height}
# Extract distinct non-NA values and generate colors dynamically (if not already run prior to p1)
msg_levels <- na.omit(unique(ans_df$Message))
msg_colors <- setNames(scales::hue_pal()(length(msg_levels)), msg_levels)

p2 <- ggplot(sd_trees, aes(x = dbh_cm, y = height, color = Message,
                           text = paste0("Tree: ", name, "<br>",
                                         "Message: ", ifelse(is.na(Message), "None", as.character(Message)), "<br>",
                                         dbh_cm, " cm \u2014 ", height, " m"))) +
  geom_point(size = 2, alpha = 0.75) +
  scale_color_manual(
    na.value = brand_green,
    values = msg_colors
  ) +
  labs(title = "Tree height vs diameter", subtitle = "Height estimated using QSM",
       x = "Diameter (cm)", y = "Height (m)", color = "Message") +
  report_theme

ggplotly(p2, tooltip = "text", height = 420) %>%
  highlight(on = "plotly_click", off = "plotly_doubleclick",
            color = accent_amber, opacityDim = 0.2) %>%
  layout(
    font = list(family = "IBM Plex Sans, sans-serif"),
    showlegend = FALSE
  )
```

## Data table

```{r allometry-table}
datatable(
  sd_trees,
  rownames = FALSE,
  filter = "top", # Will automatically use a multi-select dropdown for factors
  colnames = c(
    "Tree" = "name", 
    "Message" = "Message", 
    "Volume (m\u00B3)" = "volume",
    "Height (m)" = "height", 
    "DBH (cm)" = "dbh_cm"
  ),
  options = list(
    pageLength = 18, 
    dom = "ftip",
    columnDefs = list(
      list(
        visible = FALSE, 
        targets = which(names(ans_df) %in% c("dbh", "filename", "lon", "lat")) - 1
      )
    )
  ),
  class = "display compact stripe"
)
```

# Distributions {.tabset}

## Diameter

```{r distribution-diam}
p1 <- ggplot(ans_df, aes(x = dbh_cm,
                          text = after_stat(paste0(count, " trees, ", round(x, 1), " cm bin")))) +
  geom_histogram(bins = 30, fill = brand_green, col = "white", alpha = 0.9) +
  labs(title = "Distribution of tree diameter", subtitle = "Estimated by circle fitting",
       x = "Diameter (cm)", y = "Count") +
  report_theme

ggplotly(p1, tooltip = "text", height = 380) %>%
  layout(font = list(family = "IBM Plex Sans, sans-serif"))
```

## Height

```{r distribution-height}
p2 <- ggplot(ans_df, aes(x = height,
                          text = after_stat(paste0(count, " trees, ", round(x, 1), " m bin")))) +
  geom_histogram(bins = 30, fill = brand_green, col = "white", alpha = 0.9) +
  labs(title = "Distribution of tree height", subtitle = "Estimated using QSM",
       x = "Height (m)", y = "Count") +
  report_theme

ggplotly(p2, tooltip = "text", height = 380) %>%
  layout(font = list(family = "IBM Plex Sans, sans-serif"))
```

# Contributions

```{r contribution}
data <- ans_df[order(-ans_df$volume), ]
data$Tree <- seq_len(nrow(data))
data$Vcum <- cumsum(data$volume)
data$Vp   <- data$Vcum / sum(data$volume) * 100

half_vol_idx   <- which(data$Vp >= 50)[1]
half_vol_value <- data$Vp[half_vol_idx]

sd_contrib <- SharedData$new(data, key = ~name, group = "trees")
```

<div class="callout">
The **`r half_vol_idx`** (`r round(half_vol_idx / Ntree * 100)`%) biggest trees represent 50% of the volume.
The **`r Ntree - half_vol_idx`** (`r round((Ntree - half_vol_idx) / Ntree * 100)`%) smallest trees represent the other 50%.
</div>

Click a point below to see which tree it is (linked to the table and plots above).

```{r contribution2}
p1 <- ggplot(sd_contrib, aes(x = Tree, y = Vp,
                              text = paste0("Tree: ", name, "<br>rank ", Tree, " \u2014 ", round(Vp, 1), "% cumulative"))) +
  geom_line(color = brand_green, linewidth = 1, group = 1) +
  geom_point(color = brand_green, size = 1.6) +
  geom_hline(yintercept = 50, linetype = "dashed", color = mid_gray) +
  geom_vline(xintercept = half_vol_idx, linetype = "dashed", color = mid_gray) +
  labs(title = "Cumulative volume of trees",
       subtitle = "Trees ordered by volume, largest first",
       x = "Tree rank", y = "Cumulative volume (%)") +
  report_theme

ggplotly(p1, tooltip = "text", height = 420) %>%
  highlight(on = "plotly_click", off = "plotly_doubleclick",
            color = accent_amber, opacityDim = 0.2) %>%
  layout(font = list(family = "IBM Plex Sans, sans-serif"))
```

# Tree map

Each tree's diameter footprint is drawn at 2x scale and colored by volume. Scroll to zoom; click a tree for details — the selection stays in sync with the charts and table above.

```{r tree-map, fig.height=5}
local_xy <- st_coordinates(ans)
ans_df$x_local <- local_xy[, 1]
ans_df$y_local <- local_xy[, 2]
sd_local <- SharedData$new(ans_df, key = ~name, group = "trees")

p <- ggplot(sd_local, aes(x = x_local, y = y_local, size = volume, color = volume,
                          text = paste0("<b>Tree: ", name, "</b><br>",
                                        "Volume: ", volume, " m³<br>",
                                        "Height: ", height, " m<br>",
                                        "DBH: ", dbh_cm, " cm"))) +
  geom_point(alpha = 0.75) +
  scale_color_viridis_c(option = "plasma") +
  labs(title = "Tree map",
       x = "X", y = "Y") +
  report_theme + theme(legend.position = "right")

# Render with plotly_build to process HTML correctly
p_plotly <- ggplotly(p, tooltip = "text", height = 480) %>%
  highlight(on = "plotly_click", off = "plotly_doubleclick",
            color = accent_amber, opacityDim = 0.2)

plotly_build(p_plotly)
```

<div class="report-footer">
Generated automatically by r-lidar inc. &middot; `r format(Sys.time(), "%Y-%m-%d %H:%M")`
</div>
